### 2025-01-27 scripts/Paper-analysis.Rmd
save k-mer density expectation figure

```{r}
lr = read.csv('data/leading_region_081123.tsv', sep='\t')
plasmid.table = read.csv('data/plasmid_table.tsv', sep='\t')
retained.PTUs.lr = read.csv('data/leading-region-PTUs.txt', header=F)$V1
lr.identified.PTU.names = sort(unique(lr$PTU))
# Minimum sizes
sort(sapply(lr.identified.PTU.names, function(x) 
  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
```

@@

>>>>

```{r}
lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
lr.identified.PTU.names = sort(unique(lr$PTU))
# Minimum sizes
PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
```

@@

These PTUs are retained for analysing the leading region:

```{r}
leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
cat(leading.region.PTUs, sep="\n")
```
### Sliding window analysis

k=6 
```{r}
# Theoretical expectation for 6-bp palindrome density
theoreticalExpectation = function(p){# p is GC content
  return ((p**2 - p + 1/2)**3)
}
d = read.csv('../test_output.csv', header=F, stringsAsFactors = F)
colnames(d) = c("PTU", "plasmid", "x", "density", "GC_n")

d.reduced = d[which(d$PTU %in% leading.region.PTUs), ]

d.reduced$expected_density_new_calc = theoreticalExpectation(d.reduced$GC_n/5000)

d.reduced$diff_density = d.reduced$density-d.reduced$expected_density_new_calc

medians = d.reduced %>% group_by(PTU, plasmid) %>%
  summarise(median=median(density)) %>% group_by(PTU) %>%
  summarise(median=median(median))

# One line per plasmid
d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
  summarise(median=median(diff_density),
            n=length(diff_density), 
            median.original=median(density),
            GC=median(GC_n))

# Maximum counts
d.reduced.median.plot = d.reduced.median %>% 
  group_by(PTU) %>%
  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
  ungroup() %>%  # Remove the grouping
  select(-max_n)  # Optionally remove the max_n column if not needed
# Plot the distribution
p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median))+
  geom_hline(yintercept = 0, linetype='dashed')+
  theme_bw()+
  geom_path(aes(group=PTU, alpha=normalized_n))+
  facet_wrap(~PTU, scales="free_x", nrow=3)+
  ylab("Deviation from expected 6-mer palindrome density (5kb sliding window)")+
  labs(alpha="% of plasmids within PTU")+
  theme(panel.grid=element_blank())+  
  scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(legend.position = c(0.8, 0.2))
ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf'.
       width=10, height=8)
```

Repeat for k=4?
